---
title: "eBay cars announcements analysis"
author: "Sergey Amarin, Oleksii Zymin"
date: "1/7/2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction
<font size="2">
The goal of our analysis is to derive useful findings regarding the structure of the American cars market, it's the most popular characteristics, trends and distributions.  
We have a dataset from eBay motors website, with announcements for selling used and new cars in the USA, which we created by ourselves during the Webscapping course last semester. 
The dataset consists of around 45 000 announcements with details like:


- Brand
- Price
- Year
- Mileage
- Color
- Engine
- Body Type and etc.


The variables are both discrete and continuous, so we are able to use a wide range of visualisation technics. 
</font>

```{r, echo=FALSE, warning=FALSE, error = FALSE, message = FALSE}
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(scales)
library(tidyverse)
library(gridExtra)
library(cowplot)
library(ggforce)
library(GGally)
library(Rmisc)
library(grid)
library(pastecs)
library(dplyr)
library(treemapify)
library(qqplotr)

cars <- read.csv('D:\\UW\\2nd semester\\ML\\Project\\cars.csv')
cars <- cars[cars$Price < 100000,]
cars <- cars[cars$Mileage < 200000,]
cars <- na.omit(cars)
options(scipen=999)

cars3 <- cars %>% 
  group_by(Make) %>% 
  dplyr::summarise(number = n(), price = mean(Price))
```

## Scatter plot Price vs Year

```{r scatter, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars, aes(y=Price, x=Mileage)) + geom_point(alpha = 0.1, color = 'yellow')+geom_smooth(color = 'green')+scale_y_continuous(name = "Price in USD", breaks = seq(0, 100000, by = 10000), labels = format(seq(0, 100000, by = 10000), big.mark = ","))+scale_x_continuous(name = "Mileage in miles", breaks = seq(0, 200000, by = 50000), labels = format(seq(0, 200000, by = 50000), big.mark = " "))+ggtitle('Price vs Mileage dependency')+theme_dark()
```
The higher the Mileage - the lower the price.

## Makers tileplot

```{r tile, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(cars3, aes(area = number, label = Make, fill = price)) +
  geom_treemap() +
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre",
                    grow = F) +
  scale_fill_gradient(guide = guide_colorbar(barheight = 10))
```
The largest category 'Other' consists of a huge variety of small and medium cars brands. The top makers are Ford and Chevrolet.

## Correlation matrix

```{r corr, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
#Corrplot
corr <- cor(cars %>% select(Price, Year, Mileage, Engine, Cylinders))
ggcorrplot::ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE) 
```

<font size="2">
There isn't strong correlation between numeric variables, because cars' characteristics can be too different and individual.
However we can see one interesting finding - decreasing Cylinders number with the time. 
</font>

## Mileage boxplots

```{r box, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars, aes(y = Mileage, x = Make, fill = Make)) +
  geom_boxplot(show.legend = F) +
  coord_flip()
```

Toyota cars have the highest mean mileage, while RAM - the lowest. The mean Mileage is around 40-50 thousands miles. 

## Count hist

```{r hist, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
cars2 <- cars %>% 
  group_by(Year) %>% 
  dplyr::summarise(number = n())

Year <- cars2[cars2$Year>2000,]$Year

ggplot(data = cars2[cars2$Year>2000,], aes(x = Year, y = number)) +
  geom_bar(stat = 'identity', fill = 'darkblue', color = 'white') +
  theme_classic() +
  geom_text(aes(label = number), hjust = -.5, size = 3.5, angle = 90)+
  labs(y = "Count",  title = "Number of cars by year of prod") +
  scale_x_continuous("Year of production", labels = as.character(Year), breaks = Year)+
  scale_y_continuous(limits = c(0,7700))+
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
```

The most selling cars are 3-5 years old, because people frequently tries to get rid of new cars after warranty expiration to buy a newer model. 

## Price distribution by sellers
```{r normal, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars, aes(x = Price)) +
  geom_histogram(alpha = 0.9,position = 'stack', aes(y = ..count../sum(..count..), fill = For.Sale.By)) +
  scale_y_continuous(breaks = seq(0, .2, by = 0.01), labels = scales::percent)+theme_classic()+
  ylab('Share')+
  scale_fill_manual(values = c("Private" = "darkgrey",
                               "Dealer" = "red"
                               ))
```

And indeed, the most cars are selling by dealers.

## Normality check

```{r normal2, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
w1 <- ggplot(data = cars, aes(x = Price)) +
  geom_histogram(aes(y = ..density..)) +
  ylab('Density')+
  stat_function(fun = dnorm,
                color = 'blue',
                size = 1,
                args = list(mean = mean(cars$Price), sd(cars$Price)))
##QQ plot
w2 <- ggplot(cars, aes(sample = Price)) + 
  stat_qq_point() + # stat_gg_point instead of stat_qq()
  stat_qq_line(color = 'red') + # identity line
  stat_qq_band()

#Subplot
ggdraw() +
  draw_plot(w1 + theme(legend.justification = "bottom"),
            0, 0, 1, 1) + # size of the first graph (coordinates of bottom left corner and then height and width)
  draw_plot(w2 + theme(legend.position = 'none', 
                       axis.title = element_text(size = 9)), 
            0.5, 0.5, 0.5, 0.5) + # the same as above
  draw_plot_label(c("Normal disstibution", "QQ Plot"), c(0, 0.5), c(1, 1), size = 10)
```

The price's distribution is far away from normal.

## Price by Brand and Body Type
```{r dist, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
  ggplot(data = cars, aes(x = Price, y = ..count../sum(..count..), color = Body.Type)) +
    stat_bin(bins = 20, geom = 'line', position = 'dodge', size = 1)+
    scale_y_continuous(breaks = seq(0, 1, by = 0.01), labels = scales::percent)+
    ylab('Share')
```

Traditionally the most popular body types is USA are SUV and Pickup. However Sedan is popular as well. 

## Mean price
```{r mean, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
#Average price lineplot USA makers
cars_avg_usa <- cars[cars$Year > 1960 & cars$Make %in% c('Ram', 'Ford', 'Dodge', 'GMC', 'Chevrolet'),] %>% 
  group_by(Year) %>% 
  dplyr::summarise(avg = mean(Price), .groups = 'drop')
cars_avg_usa$Country <- 'USA'

#Average price lineplot Germany makers
cars_avg_germany <- cars[cars$Year > 1960 & cars$Make %in% c('BMW', 'Audi', 'Porsche', 'Mercedes-Benz'),] %>% 
  group_by(Year) %>% 
  dplyr::summarise(avg = mean(Price), .groups = 'drop')
cars_avg_germany$Country <- 'Germany'

#Average price lineplot Japan makers
cars_avg_japan <- cars[cars$Year > 1960 & cars$Make %in% c('Toyota', 'Lexus'),] %>% 
  group_by(Year) %>% 
  dplyr::summarise(avg = mean(Price), .groups = 'drop')
cars_avg_japan$Country <- 'Japan'

#Average price lineplot Other makers
cars_avg_other <- cars[cars$Year > 1960 & cars$Make %in% c('Other'),] %>% 
  group_by(Year) %>% 
  dplyr::summarise(avg = mean(Price), .groups = 'drop')
cars_avg_other$Country <- 'other'

cars_avg <- rbind(cars_avg_germany, cars_avg_japan, cars_avg_usa, cars_avg_other)

ggplot(data = cars_avg, aes(x = Year, y = avg, color = factor(Country))) +
  geom_line(size = 1) +
  geom_hline(yintercept=median(cars_avg$avg), linetype="dashed", 
             color = "red")+
  theme_minimal()+
  ggtitle(label = 'Average makers price by Country')+
  ylab('Price in USD')
```

Mediam cars price (red dashed) is around 25000. Germany's producers are the most expensive and Japan's are the least. Very old cars (antique) are such expensive as new ones.

## Exterior color
```{r color, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars, aes(x = Make, fill = Exterior.Color)) +
  geom_bar(position = "stack")+
  theme(axis.text.x = element_text(angle = 90, vjust = .5))+
  scale_fill_manual("legend", values = c("Black" = "black", "White" = "white", "Silver" = "lightgrey", "Red" = "red", "Green" = "green", "Gray" = "darkgray", "Blue" = "blue", "Other" = "cyan"))+
  labs(y = "Count", x = "Make", title = "Exterior colors distribution")+
  theme(panel.background = element_rect(fill = '#fffae0'))
```

The most popular exterior colors are default ones:  black, white and gray/silver.

## Transmission types share
```{r trans, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars[cars$Year>"1945",], aes(x = Year, fill = Transmission)) +
  geom_bar(position = "fill", width=1) +theme_minimal()+
  labs(y = "Share", x = "Year of production", title = "Transmission change trend")+
  scale_fill_manual(values = c("Manual" = "#353436",
                               "Automatic" = "#1b98e0"
                               ))
```

Popularity of manual transmission type is decreasing.

## Body types
```{r body, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars[sample.int(nrow(cars),100),], aes(x=Year, y=Price, shape = Body.Type, color = Fuel.Type)) + geom_point(alpha = 0.3, size = 6, stroke = 2)+scale_y_continuous(name = "Price in USD", breaks = seq(0, 100000, by = 20000), labels = format(seq(0, 100000, by = 20000), big.mark = ","))+scale_x_continuous(name = "Year", breaks = seq(1900, 2025, by = 15), labels = format(seq(1900, 2025, by = 15)))+theme_minimal()+scale_shape_manual(values=c(15, 1, 2, 17, 18, 0, 16, 3))+ggtitle('Price of various body and fuel types')
```

SUV and pickups and sedans seems to be the most popular ones nowadays, and exotic body types was popular in the past. The most vehicles are using gasoline.

## Engine Volume 
```{r mile, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
cars_engine <- cars[cars$Year >= 1950,] %>% 
  group_by(Year, Drive.Type) %>% 
  dplyr::summarise(avg = mean(Engine), .groups = 'drop')

ggplot() +
  geom_col(data = cars_engine,  aes(x = Year, y = avg, fill = Drive.Type), position = 'dodge', width = 1, alpha = 0.5)+
  ylab('Average Engine Volume')
```

AWD trucks have bigger engine volume. And it is increasing till 2000s, then fuel economy behavior appeared as well as more effective engines and engine volume trend became negative. 

## Engines boxplots
```{r box2, echo = FALSE, warning=FALSE, error = FALSE, message = FALSE}
ggplot(data = cars[cars$Year >= 2000,], aes(y = Engine, x = Drive.Type, fill = Turbo)) +
  geom_boxplot()+
  scale_fill_manual(values = c("No" = "#e016c9",
                               "Yes" = "#d9e016"
                               ))+
  theme_bw()
```

<font size="2">
From another perspective engines for both 4WD and 2WD cars have no significant differences. 2WD has more outliers with extremely high values - it could be some custom race cars. Additionally turbo cars have lower engine volume. 
</font>

## Conclusions

<font size="4">

- Cars with higher mileage is less expensive

- The highest share of American cars market is filled with american cars brands

- The most cars are 3-5 years old and selling by dealers

- SUV, Sedan and Pickup are the most popular body types.

- Germany's brands are more expensive, while Japan's are the cheapest.

- The most popular exterior colors are default ones: black, white, gray

- Manual transmission becomes less popular with time

- Engines' volumes are becoming less and less after 2000 

</font>

## Thank you!
